Structure constants and short-time simulations of open collective quantum spin systems

### 2019-03-21 soc_squeezing/collective_paper/collective_spin.tex
updated text to reflect changes to figures in plot_squeezing.py

In principle, the factorial suppression of terms at higher orders of
the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
given time $t$, there exists a truncation order $M_t$ for which the
truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
accessible with limited computational resources, such that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
allows computation of the correlator $\bk{\S_{\v m}}$ to a maximal
time $t_{M_{\t{max}}}$.  As we will see in the following section,
$M_{\t{max}}=40$ will suffice for the accurate computation of
collective spin correlators up to times that are relevant e.g.~for
spin squeezing protocols.

@@

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics
  \label{fig:squeezing_coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$
  \label{fig:squeezing_incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:squeezing_coherent} coherent and
    \sref{fig:squeezing_incoherent} incoherent evolution, computed
    using exact methods (solid lines), quantum trajectory simulations
    (dots), and the truncated Taylor expansion in
    Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed lines).
    Solid circles mark the times at which the truncated Taylor
    expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation} with $M=40$.  Squeezing is shown for
both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
as dynamics in the presence of uncorrelated emission, absorption, and
dephasing of individual spins at a rate $\chi$ (Figure
\ref{fig:squeezing_incoherent}), respectively described by the sets of
jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
appropriate transformations of the Hamiltonian and jump operators);
the only effect of this rotation on the results presented in Figure
\ref{fig:benchmarking} is to prolong the time for which the truncated
Taylor expansion agrees with the benchmarking simulations.  The reason
for different results in rotated basis has to do with the breakdown of
the truncated Taylor expansion, which we discuss further below.

@@

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, with decay
and dephasing of individual spins at rates
$\gamma_-=\gamma_\z=100\chi$.  Computations of this sort were used to
put lower bounds on theoretically achievable spin squeezing via TAT
with decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the truncated Taylor expansion and the
collective-spin structure constants calculated in this work.

\begin{figure}
  \centering
  \includegraphics{decoherence_strong.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under incoherent evolution with decay and dephasing of individual
    spins at rates $\gamma_-=\gamma_\z=100\chi$, computed using the
    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
    $M=40$.  Solid circles mark the times at which the truncated
    Taylor expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:decoherence_strong}
\end{figure}

\section{Out-of-time-ordered correlators}
\label{sec:OTOC}

Compare ``true'' OTOCs with decoherence vs.~``practical'' OTOCs in
experiments with time-reversal of coherent dynamics and decoherence

Discuss extension of this work to spin-boson systems

@@

>>>>

In principle, the factorial suppression of terms at higher orders of
the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
given time $t$, there exists a truncation order $M_t$ for which the
truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
accessible with limited computational resources, such that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
allows computation of the correlator $\bk{\S_{\v m}}$ to a maximal
time $t_{M_{\t{max}}}$.  As we will see in the following section,
$M_{\t{max}}=35$ will suffice for the accurate computation of
collective spin correlators up to times that are relevant e.g.~for
spin squeezing protocols.

@@

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics
  \label{fig:squeezing_coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$
  \label{fig:squeezing_incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:squeezing_coherent} coherent and
    \sref{fig:squeezing_incoherent} incoherent evolution, computed
    using exact methods (solid lines), quantum trajectory simulations
    (dots), and the truncated Taylor expansion in
    Eq.~\eqref{eq:order_truncation} with $M=35$ (darker dashed lines).
    Solid circles mark the times at which the truncated Taylor
    expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
as dynamics in the presence of spontaneous decay, excitation, and
dephasing of individual spins at rates $\chi$ (Figure
\ref{fig:squeezing_incoherent}), respectively described by the sets of
jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
appropriate transformations of the Hamiltonian and jump operators);
the only effect of this rotation on the results presented in Figure
\ref{fig:benchmarking} is to prolong the time for which the truncated
Taylor expansion agrees with the benchmarking simulations.  The reason
for different results in rotated basis has to do with the breakdown of
the truncated Taylor expansion, which we discuss further below.

@@

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
spontaneous decay, excitation, and dephasing of individual spins at
rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computations of this
sort were used to put lower bounds on theoretically achievable spin
squeezing via TAT with decoherence in
Ref.~[\citenum{he2019engineering}], exemplifying a concrete
application of the truncated Taylor expansion and the collective-spin
structure constants calculated in this work.

\begin{figure}
  \centering
  \includegraphics{decoherence_strong.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under incoherent evolution spontaneous with decay, excitation, and
    dephasing of individual spins at rates
    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the
    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
    $M=35$.  Solid circles mark the times at which the truncated
    Taylor expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:decoherence_strong}
\end{figure}

\section{Two-time correlation functions}
\label{sec:TTC}

These are very easy to compute.

% \section{Out-of-time-ordered correlators}
% \label{sec:OTOC}

% Compute some OTOCs?  There is a small chance that doing this will
% take a prohibitively long amount of time...

% Possibility: compare ``true'' OTOCs with decoherence
% vs.~``practical'' OTOCs in experiments with decoherence and
% time-reversal of coherent dynamics.

@@

Mention/discuss extension of this work to spin-boson systems.
